TO FIND

### 2013-05-28 ms.tex
Figure legends and formatting

% TODO:
% DG: First complete reading
% DG: Figures
% DG: Legends
% DG: Compile references
% TP & PDP: fix bugs I generated!!!
% DG: Abstract

@@

2: Qu\'ebec Centre for Biodiversity Sciences, Stewart Biological Sciences Building, 1205 Dr.~Penfield Avenue, Montr\'eal (QC), H3A 1B1, Canada\\

\textbf{Words in the abstract:}      \\
\textbf{Words in the main text:}    \\
\textbf{Words in the legends:}    \\
\textbf{References:}             \\
\textbf{Table:}                    \\

@@

%========================================================%
\section{Introduction}

@@

At first sight, the development of neutral theory appears as a step back.
Neutral theory of biodiversity makes the provocative assumption that species are
ecologically equivalent \parencite{Bell2000,Hubbell2001}. Neutral ecological communities are
driven only by demographic stochasticity and dispersa and thus, variation in the environment has no impact on
demography. Neutral theory sparked
an historical debate still lasting after more than a decade
\parencite{Chave2004, Etiennee2011, Rosindell2012,Clark2012}. It was stimulated
by the surprising ability of neutral models to fit some well studied empirical
observations such as species abundance distributions and distance-decay
relationships.

@@

Our main objective is to use neutral theory to stress the importance of
landscape network structure on the distribution of diversity. We refer to the
landscape organization as a "spatial contingency" \parencite{Peres-Neto2013}
that could potentially affect the coexistence mechanisms at play. We will
therefore move from a perspective where dispersal is either global or spatially
explicit (e.g. over a lattice), and spatial constant, to a perspective focusing
on the variance of dispersal. A second generation of neutral models (e.g.
\textcite{Economo2008, Economo2011,Desjardins2012a,Desjardins2012b}, and even
experiments \parencite{Altermat2012}, recently introduced more realistic
landscapes and found surprising contributions of spatial contingencies. We start
with a short review of the main approaches to describe spatial networks. Then we
describe three simple toy models of metacommunity dynamics, using this
opportunity to review their assumptions and main predictions. We provide as
Supplementary Material the R scripts for the toy models and all simulations
conducted for this paper. We then conduct simple simulations of these models to
reveal with simple examples the impact of spatial network structure on diversity
distribution. We conclude with a discussion on the operationally of the
framework.

%========================================================%
\section{Network representation of landscapes}

@@

\begin{table*}
	\centering
	\begin{tabular}{p{5cm}p{8cm}}
		\hline
		\textbf{Concept} & \textbf{Definition} \\
		\hline
		Path & A sequence of edges forming a sequence of nodes \\
		Connection & Two nodes are connected if there is a path between them \\
		Euclidean distance & Geographical distance between two nodes \\
		Geodesic distance & Length of the shortest path between two nodes \\

		\textbf{Node-level metrics} & \\
		Degree & The number of edges of a node \\
		Closeness centrality & Average geodesic distance between a node and all other vertex \\
		Eigenvector centrality & A measure of centrality based on the concept that connection to
		highly connected nodes are more important \\
		Betweenness centrality & The number of shortest paths from all nodes to all others that
		pass through that note \\
		\hline
	\end{tabular}
	\caption{Main descriptors of spatial networks used in this study.}
\end{table*}

@@

\begin{table*}
	\centering
	\begin{tabular}{llccc}
	\hline
	& Definition & Patch dynamics & Neutral & Species-sorting \\
	\hline
	\textbf{Variables} &     &    &    &   \\
	$p$ & Occupancy  & X &    &   \\
	$N$ & Local population size &    & X & X \\
	$Z$ & Local rel. abund. &  & X & X \\
	$P$ & Rel. abund. in the neighborhood &  & X & X \\
	$s$ & Local species richness & X & X & X \\
	$d$ & Node degree & X & X & X \\
	$C$ & Prob. of a colonization event & X &  &  \\
	$I$ & Prob. of a colonization event & X &  &  \\
	$Pr$ & Recruitment prob. &  & X & X \\
	$\lambda$ & Survival prob. &  &  & X \\

	\textbf{Parameters} & & & &\\
	$S$ & Size of regional species pool & X & X & X \\
	$c$ & Colonization prob. & X & & \\
	$e$ & Extinction prob. & X & & \\
	$J$ & Local carrying capacity & & X & X \\
	$m$ & immigration prob. from neigh. & & X & X \\
	$M$ & immigration prob. from metaco. & & X & X \\
	$k$ & Death prob. & & X & X \\
	$u$ & Niche optimum & & & X \\
	$b$ & Niche breadth & & & X \\
	$E$ & Microsite env. conditions & & & X \\
	$\overline{E}$ & Local env. average & & & X \\
	$\sigma$ & Local env. variance & & & X \\
	$\overline{E_R}$ & Regioal env. average & & & X \\
	$\sigma_R$ & Regional env. variance & & & X \\
	\hline
	\end{tabular}
	\caption{List of variables, indices and parameters from the three models}
\end{table*}

@@

\section{Results}

@@

%========================================================%
\section{Discussion}

Our objective in this paper was to review the main assumptions of three
metacommunity models and illustrate how the implementation of more realistic
landscapes could reveal the importance of dispersal on community structure. We
argued in the introduction that neutral theory is useful both to understand and
predict the impact of dispersal on community organization. The review of the
different models shows that the fundamental difference between a neutral model
and the patch dynamics model is the effect of competitive interactions on
distribution, while the difference between the neutral and the species sorting
models is the effect of unequal competitive interactions. The neutral model is
thus a usefull tool to understand the joint effects of dispersal and community
interactions. Our comparison of the distribution of $\alpha$ diversity was
particularly meaningful in that respect. The simulation results show that
competitive interactions buffer the potential impact of landscape structure. The
strongest centrality-species richness relationship was observed for the patch
dynamics, a model without any interactions. On the other hand, strong and
unequal competitive interactions minimized the effect of centrality. Our model
analysis greatly illustrates the growing recognition in metacommunity ecology
that we must move toward more realistic landscapes \parencite{Gilarranz2012}.
For field ecologists, and particularly limnologists, our review emphasizes that
we need to go beyond geographic based analysis of $\beta$ diversity (e.g.
\parencite{Legendre2005}) to topological based analyses
\parencite{Peterson2013}.

@@

The multivariate variance partitioning framework originally proposed by
\textcite{Borcard1992} and further developed by \textcite{Borcard2002} has been
widely used to quantify the relative importance of species sorting and dispersal
limitations in species distribution. This framework was originally proposed to
model species distribution as a function of environmental variables, taking into
account the spatial autocorrelation of species distribution
\parencite{Leduc1992, Borcard1992, Legendre1993}. This methodology has been
widely used over the last decade as a test of the neutral theory, its underlying
assumption and a quantification of dispersal limitations (e.g.
\textcite{Svenning2004; Hardy2004; Gilbert2004; Cottenie2005}. One problem of
this approach is however that it makes a weak test of neutrality
\parencite{McGill200}, based on the description of spatial community structure,
rather than hypothesis testing. The different models we reviewed in this article
could be better employed if used to generate null expectations of species
distribution based on different hypotheses and then compare them. But most of
all, parametrized spatially explicit neutral models could be more useful if used
to predict biodiversity under different global change scenarios. For instance,
neutral models could be used to predict the consequences of habitat destruction,
fragmentation or a change in the connectivity matrix \parencite{Hubbell2008}.
The spatially explicit description of the landscape is a major improvement
toward that end, providing much flexibility in the scenarios that could be
explored.

@@

%========================================================%
\section{Acknowledgements}
DG received financial support from NSERC and Canada Research Chair program. TP
is funded by a MELS-FQRNT post-doctoral fellowship and PDP by a NSERC
fellowship.
\newpage

@@

\newpage
%========================================================%
\section*{Figure legends}

%------------------------
\subsection*{Figure 1}
\textbf{Illustration of the four simulated landscapes}. The color code
represents the local species richness simulated with a neutral model, ranked
from the poorest (red) to the richest (blue). Parameters are: $P = 100$, $N =
25$, $m = 0.2$, $M = 0.01$, $k = 0.1$, $J_x = 100$.

%------------------------
\subsection*{Figure 2}
\textbf{Relationship between local species richness and metrics of node
centrality}. The upper two panels are simulation results conducted with the
random geometric graph illustrated at Fig. 1 and the lower two panels are runs
with the random tree graph.

%------------------------
\subsection*{Figure 3}
\textbf{Bray curtis dissimilarity as a function of geographic distance}.

%------------------------
\subsection*{Figure 4}
\textbf{Bray curtis dissimilarity as a function of topological distance}.

\newpage
%========================================================%

\section*{Figures}

%------------------------
\subsection*{Figure 1}

@@

>>>>

% TODO:
% DG: Finish figures
% DG: Compile references
% TP & PDP: fix bugs I generated!!!

@@

2: Qu\'ebec Centre for Biodiversity Sciences\\

\textbf{Keywords:} metacommunity; neutral theory; species sorting; metapopulation; spatial network; centrality \\

\textbf{Words in the abstract:}

\textbf{Words in the main text:}

\textbf{Words in the legends:}

\textbf{Figures:} 4

\textbf{Tables:} 2

\textbf{References:}

@@

%========================================================%
\section*{Abstract}
The metacommunity perspective appears naturally as the appropriate conceptual
framework to make ecology more predictive, integrating elements of theoretical
ecology. The recent development of neutral theory appears as a step back in that
direction because of the assumption of ecological equivalence and the absence of
any effect of the environment on community organization. A remarkable strength
of neutral theory is nonetheless to provide a general theory of diversity that
accounts for a wide range of empirical observations. In this paper, we argue
that neutral theory can be a useful tool to understand the impact of dispersal
on community organization in landscapes of various complexities. Our main
objective is to use neutral theory to stress the importance of complex landscape
network structure on the distribution of diversity. We refer to the landscape
organization as a "spatial contingency" that could potentially affect the
coexistence mechanisms at play. We briefly review the main approaches to
describe spatial networks and describe three simple toy models of metacommunity
dynamics. We take this opportunity to review their assumptions and main
predictions. We then conduct simple simulations of these models to reveal with
simple examples the impact of spatial network structure on diversity
distribution. The simulation results show that competitive interactions buffer
the potential impact of landscape structure. The strongest centrality-species
richness relationship was observed for the patch dynamics, a model without any
interactions. On the other hand, strong and unequal competitive interactions
minimized the effect of centrality. We conclude that the neutral model is thus a
useful tool to understand the joint effects of dispersal and community
interactions. Our analysis shows that ecologists must now integrate more
realistic landscapes when analyzing community assembly from a metacommunity
perspective.
\newpage

%========================================================%
\section*{Introduction}

@@

At first sight, the development of neutral theory appears as a step back.
Neutral theory of biodiversity makes the provocative assumption that species are
ecologically equivalent \parencite{Bell2000,Hubbell2001}. Neutral ecological
communities are driven only by demographic stochasticity and dispersa and thus,
variation in the environment has no impact on demography. Neutral theory sparked
an historical debate still lasting after more than a decade
\parencite{Chave2004, Etienne2011, Rosindell2012,Clark2012}. It was stimulated
by the surprising ability of neutral models to fit some well studied empirical
observations such as species abundance distributions and distance-decay
relationships.

@@

Our main objective is to use neutral theory to stress the importance of
landscape network structure on the distribution of diversity. We refer to the
landscape organization as a \emph{spatial contingency} \parencite{Peres-Neto2013}
that could potentially affect the coexistence mechanisms at play. We will
therefore move from a perspective where dispersal is either global or spatially
explicit (e.g. over a lattice), and spatial constant, to a perspective focusing
on the variance of dispersal. A second generation of neutral models (e.g.
\textcite{Economo2008, Economo2011,Desjardins2012a,Desjardins2012b}, and even
experiments \parencite{Altermat2012}, recently introduced more realistic
landscapes and found surprising contributions of spatial contingencies. We start
with a short review of the main approaches to describe spatial networks. Then we
describe three simple toy models of metacommunity dynamics, using this
opportunity to review their assumptions and main predictions. We provide as
Supplementary Material the R scripts for the toy models and all simulations
conducted for this paper. We then conduct simple simulations of these models to
reveal with simple examples the impact of spatial network structure on diversity
distribution. We conclude with a discussion on the operationally of the
framework.

%========================================================%
\section*{Network representation of landscapes}

@@

@@

@@

\section*{Results}

@@

%========================================================%
\section*{Discussion}

Our objective in this paper was to review the main assumptions of three
metacommunity models and illustrate how the implementation of more realistic
landscapes could reveal the importance of dispersal on community structure. We
argued in the introduction that neutral theory is useful both to understand and
predict the impact of dispersal on community organization. The review of the
different models shows that the fundamental difference between a neutral model
and the patch dynamics model is the effect of competitive interactions on
distribution, while the difference between the neutral and the species sorting
models is the effect of unequal competitive interactions. The neutral model is
thus a useful tool to understand the joint effects of dispersal and community
interactions. Our comparison of the distribution of $\alpha$ diversity was
particularly meaningful in that respect. The simulation results show that
competitive interactions buffer the potential impact of landscape structure. The
strongest centrality-species richness relationship was observed for the patch
dynamics, a model without any interactions. On the other hand, strong and
unequal competitive interactions minimized the effect of centrality. Our model
analysis greatly illustrates the growing recognition in metacommunity ecology
that we must move toward more realistic landscapes \parencite{Gilarranz2012}.
For field ecologists, and particularly limnologists, our review emphasizes that
we need to go beyond geographic based analysis of $\beta$ diversity (e.g.
\parencite{Legendre2005}) to topological based analyses
\parencite{Peterson2013}.

@@

The multivariate variance partitioning framework originally proposed by
\textcite{Borcard1992} and further developed by \textcite{Borcard2002} has been
widely used to quantify the relative importance of species sorting and dispersal
limitations in species distribution. This framework was originally proposed to
model species distribution as a function of environmental variables, taking into
account the spatial autocorrelation of species distribution
\parencite{Leduc1992, Borcard1992, Legendre1993}. This methodology has been
widely used over the last decade as a test of the neutral theory, its underlying
assumption and a quantification of dispersal limitations (e.g.
\textcite{Svenning2004, Hardy2004, Gilbert2004, Cottenie2005}. One problem of
this approach is however that it makes a weak test of neutrality
\parencite{McGill200}, based on the description of spatial community structure,
rather than hypothesis testing. The different models we reviewed in this article
could be better employed if used to generate null expectations of species
distribution based on different hypotheses and then compare them. But most of
all, parametrized spatially explicit neutral models could be more useful if used
to predict biodiversity under different global change scenarios. For instance,
neutral models could be used to predict the consequences of habitat destruction,
fragmentation or a change in the connectivity matrix \parencite{Hubbell2008}.
The spatially explicit description of the landscape is a major improvement
toward that end, providing much flexibility in the scenarios that could be
explored.

@@

%========================================================%
\section*{Acknowledgements}
DG received financial support from NSERC and Canada Research Chair program. TP
is supported by a MELS-FQRNT post-doctoral fellowship and PDP by a NSERC
fellowship.
\newpage

@@

%========================================================%
\section*{Figure legends}

%------------------------
\subsection*{Figure 1}
\textbf{Illustration of the four simulated landscapes}. The color code
represents the $\alpha$ diversity simulated with a neutral model, ranked from
the poorest (red) to the richest (blue). Parameters: $N = 25$, $r = 0.3 $, $S =
100$, $m = 0.2$, $M = 0.01$, $k = 0.1$, $J_x = 100$. Simulations run 1000 time
steps.

%------------------------
\subsection*{Figure 2}
\textbf{Relationship between $\alpha$ diversity and node centrality}. The upper
two panels are simulation results conducted with the random geometric graph
illustrated at Fig. 1 and the lower two panels are runs with the random tree
graph. Parameters:$S = 100$, $c = 0.4$, $e = 0.1$, $J_x = 100$, $m = 0.2$, $M =
0.01$, $k = 0.1$, $ u \epsilon [0,100]$, $b = 15$, $ E_x \epsilon [0,100]$,
$\overline{E_R} = 50$, $\sigma{E_R} = 5$. Simulations run 1000 time steps

%------------------------
\subsection*{Figure 3}
\textbf{Bray curtis dissimilarity as a function of Euclidean distance}. Parameters as in Fig. 2.

%------------------------
\subsection*{Figure 4}
\textbf{Bray curtis dissimilarity as a function of geodesic distance}. Parameters as in Fig. 2.

\newpage
%========================================================%
%------------------------
\subsection*{Figure 1}

@@

%========================================================%

\begin{table*}[c]
	\centering
	\begin{tabular}{p{5cm}p{8cm}}
		\hline
		\textbf{Concept} & \textbf{Definition} \\
		\hline
		Path & A sequence of edges forming a sequence of nodes \\
		Connection & Two nodes are connected if there is a path between them \\
		Euclidean distance & Geographical distance between two nodes \\
		Geodesic distance & Length of the shortest path between two nodes \\

		\textbf{Node-level metrics} & \\
		Degree & The number of edges of a node \\
		Closeness centrality & Average geodesic distance between a node and all other vertex \\
		Eigenvector centrality & A measure of centrality based on the concept that connection to
		highly connected nodes are more important \\
		Betweenness centrality & The number of shortest paths from all nodes to all others that
		pass through that note \\
		\hline
	\end{tabular}
	\caption{Main descriptors of spatial networks used in this study.}
\end{table*}
\newpage

%------------------------

\begin{table*}[c]
	\centering
	\begin{tabular}{llccc}
	\hline
	& Definition & Patch dynamics & Neutral & Species-sorting \\
	\hline
	\textbf{Variables} &     &    &    &   \\
	$p$ & Occupancy  & X &    &   \\
	$N$ & Local population size &    & X & X \\
	$Z$ & Local rel. abund. &  & X & X \\
	$P$ & Rel. abund. in the neighborhood &  & X & X \\
	$s$ & Local species richness & X & X & X \\
	$d$ & Node degree & X & X & X \\
	$C$ & Prob. of a colonization event & X &  &  \\
	$I$ & Prob. of a colonization event & X &  &  \\
	$Pr$ & Recruitment prob. &  & X & X \\
	$\lambda$ & Survival prob. &  &  & X \\

	\textbf{Parameters} & & & &\\
	$S$ & Size of regional species pool & X & X & X \\
	$c$ & Colonization prob. & X & & \\
	$e$ & Extinction prob. & X & & \\
	$J$ & Local carrying capacity & & X & X \\
	$m$ & immigration prob. from neigh. & & X & X \\
	$M$ & immigration prob. from metaco. & & X & X \\
	$k$ & Death prob. & & X & X \\
	$u$ & Niche optimum & & & X \\
	$b$ & Niche breadth & & & X \\
	$E$ & Microsite env. conditions & & & X \\
	$\overline{E}$ & Local env. average & & & X \\
	$\sigma$ & Local env. variance & & & X \\
	$\overline{E_R}$ & Regioal env. average & & & X \\
	$\sigma_R$ & Regional env. variance & & & X \\
	\hline
	\end{tabular}
	\caption{List of variables, indices and parameters from the three models}
\end{table*}
\newpage
